### 2021-10-22 plos_latex_template.tex
Some edits by LA added by AST

\newcommand{\fede}[1]{\textcolor{ForestGreen}{Fede: #1}}
\newcommand{\rocio}[1]{\textcolor{MidnightBlue}{Rocio: #1}}
\newcommand{\doro}[1]{\textcolor{RedViolet}{Dorothea: #1}}
\newcommand{\jani}[1]{\textcolor{YellowOrange}{Janina: #1}}
\newcommand{\nextcoauthor}[1]{\textcolor{RawSienna}{helloauthors: #1}}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules towards an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Name1 Surname\textsuperscript{1,2\Yinyang},
Name2 Surname\textsuperscript{2\Yinyang},
Name3 Surname\textsuperscript{2,3\textcurrency},
Name4 Surname\textsuperscript{2},
Name5 Surname\textsuperscript{2\ddag},
Name6 Surname\textsuperscript{2\ddag},
Name7 Surname\textsuperscript{1,2,3*},
with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

@@

The authors of this article participated in the organization of the annual user conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and arranged by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.\fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre} \doro{remove or, replace with -} or `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, \doro{add "sharing and"} producing knowledge, and becoming active community members that conferences can provide to the non-marginalized \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can \doro{Suggestion: remove "even"} even divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion.

% Rocío: Main message of the paragraph: Roadmap of the paper
This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for an online component. \doro{stronger: "to have a strong online component for all conferences"}
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for everyone \doro{is everyone too bold? everyone is never happy. "many"?}. 
Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Rules 1, 2 and 10 refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The next two rules are focused on the people that participate in the conference: 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
Rules 5 to 9 are rules about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. 
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it:
Will a more diverse conference translate into an even gender distribution in your speakers? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
\doro{I would leave out the following sentence. It doesn't add a new thought and only partially repeats the sentence before. The questions are a stronger argument. Or we could write: "The answer to these questions depend on your field, region or community."}
Depending on the field, region, or community, it could be Black people, LGBTQIA+, Muslim, or other people.
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account. 
% This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way.

@@

% Rocío: main idea: Code of conduct for a safe place
Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
%Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.

% Rocío: Could have a paragraph about active bystandership—Andrea

@@

% Rocío: Idea of this paragraph: build a diverse team
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort and pay special attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). 
To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. 
If you already have an organizing team, check for gaps in its composition.

@@

% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging (see \textbf{Rule \ref{rule_inclusion}}).

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. 
Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly, offer leading positions, and let people from privileged groups step down. 
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
Offer support and encouragement if you encounter cases of impostor syndrome. 
You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.

@@

\section{Consciously unbias your spotlight roles}
\label{rule_unbias}

@@

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
In academia and the industry, not only is the list of people considered in spotlight roles usually biased, but the spotlight roles themselves are.
Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. \doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
Alternatively, a conference could have a hybrid format with an in-person and an online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020} and should not be dealt with at the last minute. Key decisions in this respect are hard to correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs. 
\doro{repetition with the next sentence}
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free options.

@@

Accessibility practices must include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to a broader spectrum of people. For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communications of the conference (see \textbf{Rule \ref{rule_communication}}).

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international conferences, the linguistic diversity of conference participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker, a privilege.
Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
and conferences may miss potentially innovative contributions.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
(see https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example of conversion rates based on country of origin and career status). 
In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}. 
Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
Scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and conferences usually ask for cover letters or applications from them. 
These programs are important for in-person participation, which is more expensive than online participation, to support travel and lodging expenses. 
For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper. 
Lower costs for attendance and offering financial support instead of prestigious grants--that could look good in a CV--may reduce the number of people applying for them and allow you to help all applicants. 
Applying for loans, grants, and scholarships might be an emotionally demanding task, so if this scenario is likely, simplify the process of asking for financial support. 
%Dorothea: I understand what we want to say, but there is a slight contradiction between "reduce the prestige so that less people apply and we can help everyone and reduce the barriers so that more people apply"

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR! 2021 and supporting us through the process.

>>>>

\newcommand{\fede}[1]{\textcolor{ForestGreen}{Fede: #1}}
\newcommand{\rocio}[1]{\textcolor{MidnightBlue}{Rocio: #1}}
\newcommand{\doro}[1]{\textcolor{RedViolet}{Dorothea: #1}}
\newcommand{\jani}[1]{\textcolor{YellowOrange}{Janina: #1}}
\newcommand{\la}[1]{\textcolor{RawSienna}{Laura: #1}}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules towards hosting an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Name1 Surname\textsuperscript{1,2\Yinyang},
Name2 Surname\textsuperscript{2\Yinyang},
Name3 Surname\textsuperscript{2,3\textcurrency},
Name4 Surname\textsuperscript{2},
Name5 Surname\textsuperscript{2\ddag},
Name6 Surname\textsuperscript{2\ddag},
Name7 Surname\textsuperscript{1,2,3*},
with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

@@

The authors of this article participated in the organization of the annual user conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level. The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events. \fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion. %https://www.cscce.org/resources/organizing-community-events/

% Rocío: Main message of the paragraph: Roadmap of the paper
This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
The third group includes rules 5 to 9. These rules are about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to academic, industry, or mixed conferences. While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults centered around which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
%"The answer to these questions depend on your field, region or community.
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.

@@

% Rocío: main idea: Code of conduct for a safe place
Adopting a code of conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
%Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The code of conduct team must receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the code of conduct teamwork.
% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.

% Rocío: Could have a paragraph about active bystandership—Andrea
%ast: tiny: check CoC capitalization (LA)

@@

% Rocío: Idea of this paragraph: build a diverse team
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort and pay special attention to to intersectional diversity (see \textbf{Rule \ref{rule_diversity}}). 
To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the general population.
If you already have an organizing team, check for gaps in its composition.

@@

% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. 
Do not expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Instead, nominate directly and offer leading positions, and let people from privileged groups step down.
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
Offer support and encouragement if you encounter cases of impostor syndrome. 
You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.

@@

\section{Consciously mitigate bias in your spotlight roles}
\label{rule_unbias}

@@

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
In academia and the industry, not only is the list of people considered in spotlight roles usually biased, but the spotlight roles themselves are.
Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. \doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest and offer them space to showcase them.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well into a hybrid format that maximizes the experience of everyone attending your conference. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct. 
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free, and alcohol-free options.

@@

Accessibility practices must include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communications of the conference (see \textbf{Rule \ref{rule_communication}}).

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international conferences, the linguistic diversity of conference participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker, a privilege.
Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
and conferences may miss innovative contributions.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}.
(see https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example of conversion rates based on country of origin and career status). 
In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}. 
Offering fee waivers to attendees is also a good option, but even then, other costs can be prohibitive. 
Scholarships or grants to attend the conference is a common way to boost participation from people from marginalized groups, and conferences usually ask for cover letters or applications from them. 
These programs are important for in-person participation to support travel and lodging expenses. 
For online conferences, granting fee waivers or support for hidden costs like internet connection is easier and cheaper. 
Lower costs for attendance and offering financial support instead of prestigious grants--that could look good in a CV--may reduce the number of people applying for them and allow you to help all applicants. 
Applying for loans, grants, and scholarships might be an emotionally demanding task, so if this scenario is likely, simplify the process of asking for financial support. 
%Dorothea: I understand what we want to say, but there is a slight contradiction between "reduce the prestige so that less people apply and we can help everyone and reduce the barriers so that more people apply"

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process.
